---
title: "Resistome and *Escherichia coli* strains in pediatric fecal samples"
description: |
  Report of findings from metagenome data.
author:
  - name: Sara G. Cifuentes1, Jay Graham2, Fernanda Loayza1, Carlos Saraiva1, Liseth Salinas1, Gabriel Trueba1 , Paúl A. Cárdenas1
    affiliation: 1Universidad San Francisco de Quito USFQ, Colegio de Ciencias Biológicas y Ambientales, Instituto de Microbiología, Quito, Pichincha, Ecuador; 2University of California, Environmental Health Sciences Division, Berkeley, California, USA

    affiliation_url: https://www.usfq.edu.ec/es/institutos-de-investigacion/instituto-de-microbiologia; https://publichealth.berkeley.edu/
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    theme: journal
    fig_height: 6
    fig_width: 12
    fig_caption: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 10)
```

# ***Escherichia coli* strain** detection and characterization on fecal samples

The method used to identify and characterize ***E. coli* strains** in fecal samples was taken and modified from [PanPhlAn 3.](https://github.com/SegataLab/panphlan)

> *"PanPhlAn is a strain-level metagenomic profiling tool for identifying the gene composition of individual strains in metagenomic samples."*^[Integrating taxonomic, functional, and strain-level profiling of diverse microbial communities with bioBakery 3 Francesco Beghini, Lauren J McIver, Aitor Blanco-Miguez, Leonard Dubois, Francesco Asnicar, Sagun Maharjan, Ana Mailyan, Andrew Maltez Thomas, Paolo Manghi, Matthias Scholz, Mireia Valles-Colomer, George Weingart, Yancong Zhang, Moreno Zolfo, Curtis Huttenhower, Eric A Franzosa, Nicola Segata. eLife (2021)] PanPhlAn is a tool for culture-free microbial population studies based on metagenomic data and it does not require genome assembly. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

```

## Graphs

```{r, include=FALSE}
library(ComplexHeatmap)
library(tidyverse)
library(readr)
library(gridtext)
library(dplyr)
library(base)
```

```{r, include=FALSE, echo=FALSE}
profile_ecoli_df <- read_csv("matrix_profile_ecoli")

ecoli_pangenome_annot <- read.delim("panphlan_Escherichia_coli_annot.tsv", header=FALSE)

colnames(ecoli_pangenome_annot)[1] <- "Annotation"

colnames(ecoli_pangenome_annot) [2] <- "Genes"

ecoli_pangenome_annot <- ecoli_pangenome_annot %>% select(Annotation:Genes)

nodup_ecoli_pangenome <-  ecoli_pangenome_annot %>% 
  distinct(Annotation, .keep_all = TRUE)

```

```{r, include=FALSE, echo=FALSE}
test1 <- profile_ecoli_df %>% 
  head(100) %>% 
  rowwise() %>% 
  rename_with(~gsub("_ecoli.csv", "", .x)) %>% 
  left_join(nodup_ecoli_pangenome) %>% 
  relocate(Genes, .before = PR002) %>% 
  select(-Annotation) %>% 
  remove_rownames() %>% 
  column_to_rownames(var = 'Genes') %>% 
  select((1:50), -PR019, -PR040, -PR044, -PR051, -PR071, -PR125, -PR126)

```

## Heatmap and hierarchical clustering analysis

```{r, echo=FALSE, warning=FALSE}
colors <- c("yellow", "navyblue")

Heatmap(test1,
        name = "presence = 1/absence = 0",
        column_title = gt_render ("High diversity of <span style = 'color: navyblue'>*Escherichia coli*</span> strains<br>in pediatric fecal samples in rural and peri-urban areas of Quito, Ecuador"),
        column_split = 3, 
        column_title_gp = gpar(fontsize = 10, fontface = "bold"),
        row_title = "Genes",
        row_split = 3,
        row_names_gp = gpar(fontsize = 3),
        column_names_gp = gpar(fontsize = 5), 
        col = colors,
        column_dend_height = unit(1, "cm"),
        clustering_distance_columns = "euclidean",
        clustering_method_columns = "single", 
        clustering_distance_rows = "euclidean",
        heatmap_height = unit (10, "cm")) 

```

```{r, include=FALSE, echo=FALSE}
#Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.
```






